---
title: "Assignment01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1

## Read in the data

```{r packages}
library(data.table)
library(tidyverse)
library(leaflet)
```

```{r data-read, cache=TRUE}
data2019 <- data.table::fread("ad_viz_plotval_data_2019.csv")
data2004 <- data.table::fread("ad_viz_plotval_data_2004.csv")
```

## Check the dimensions, headers, footers, variables

```{r}
dim(data2004)

head(data2004)

tail(data2004)

str(data2004)

dim(data2019)

head(data2019)

tail(data2019)

str(data2019)
```

## Check for the key variable we are analyzing

```{r}
table(data2004$ Date)
table(data2004$ "Daily Mean PM2.5 Concentration")
summary(data2004$ Date)
summary(data2004$ "Daily Mean PM2.5 Concentration")

table(data2019$ Date)
table(data2019$ "Daily Mean PM2.5 Concentration")
summary(data2019$ Date)
summary(data2019$ "Daily Mean PM2.5 Concentration")
```

Summary: According to the summary function output, 

## Step 2

## Combine the two years of data into one data frame and change the names of the key variables

```{r}
dataPM2.5 <- rbind(data2004, data2019)

dataPM2.5$Year <- format(dataPM2.5$Time, format="%Y")

rename(dataPM2.5, PM2.5 = "Daily Mean PM2.5 Concentration")
rename(dataPM2.5, site = "Site Name")
rename(dataPM2.5, lat = "SITE_LATITUDE")
rename(dataPM2.5, lon = "SITE_LONGITUDE")
```

## Step 3

## Create a basic map in leaflet

```{r}
Year.pal <- colorNumeric(c('darkgreen','goldenrod'), domain=dataPM2.5$Year)
Year.pal

map <- leaflet(dataPM2.5) %>%
  addProviderTiles('SiteMap') %>% 
  addCircles(., lat=~lat,lng=~lon, opacity=1, color = ~ Year.pal(Year), fillOpacity=1, radius=100) %>% 
```

## Step 4

## Check for any missing or implausible values

```{r}
which(is.na(dataPM2.5$PM2.5))
```

## Step 5

## Create exploratory plots

```{r}
dataPM2.5[!is.na(PM2.5)] %>% 
  ggplot()+
  geom_boxplot(mapping=aes(x=Year, y=PM2.5, fill=site))
  
ggplot(dataPM2.5) + 
  geom_histogram(mapping = aes(x = PM2.5))

ggplot(data = dataPM2.5) + 
  geom_smooth(mapping = aes(x = PM2.5, y = rh, linetype = region))
```











































