---
title: "Lab05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(data.table)
library(tidyverse)
```

```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
```

```{r}
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

```{r merging}
met <- merge(
  x = met,
  y = stations,
  all.x = TRUE, all.y = FALSE,
  by.x = "USAFID", by.y = "USAF"
) 
```
## Question 1
First, generate a representative version of each station. We well use the averages (median could also be a good way to represent it, but it will depend on the case).

```{r}
station_averages <- met[,.(
  temp      = mean(temp, na.rm =TRUE),
  wind.sp   = mean(wind.sp, na.rm =TRUE),
  atm.press = mean(atm.press, na.rm =TRUE),
  STATE = unique(STATE)
), by = USAFID]
```

Now, we need to identify the median per variable

```{r}
station_averages[,.(
  temp_50      = quantile(temp, probs = .5, na.rm = TRUE),
  wind.sp_50   = quantile(wind.sp, probs = .5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = .5, na.rm = TRUE)
)]

medians
```

Now we can find the stations that are the closest to these. (hint: `which.min()`)

```{r}
station_averages[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- station_averages[order(temp_dist)][1]
median_temp_station
```

The median temperature station is `r median_temp_station$USAFID`

## Question 2

```{r}
merge(x = station_averages, y = statioins, by.x = "USAFID", by.y = "USAF", all.x = TRUE, all.y = FALSE)
```

```{r}
station_averages[, temp_50 := quantile(temp, probs = .5, na.rm = TRUE), by = STATE]
station_averages[, wind.sp_50 := quantile(wind.sp, probs = .5, na.rm = TRUE), by = STATE]
```

Now the euclidean distance...$\sqrt{\sum_i(x_i - y_i)^2}$

```{r}
station_averages[, eudist := sqrt(
  +
)]
```


```{r}
met[, state_temp := mean(temp, na.rm = TRUE), by = STATE]
met[,temp_cat := fifelse(
  state_temp < 20, "low_temp",
  fifelse(state_temp < 25, "mid_temp", "high_temp"))
  ]
```

Now, let's make sure that we don't have NAs

```{r}
table(met$temp_cat, useNA = "always")
```

```{r}
met[, .(
  N_entries = .N,
  N_stations - unique(USCFID)
),by = temp_cat
    ]
```











